---
title: "BioCro II Paper: Section 1.1 Example"
output:
  html_document:
    toc: true
    toc_depth: 2

vignette: >
  %\VignetteIndexEntry{BioCro II Paper: Section 1.1 Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

bibliography: references/BioCro.bibtex
---
```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  error = FALSE,
  collapse = TRUE,
  cache = TRUE,
  autodep = TRUE
)
```

<!-- Note that we start with a level-2 heading so that, following best
practices, the document title is the one and only h1 element in the
document. -->

## Introduction

This is a demonstration of the example discussed in Section 1.1 and
Appendix 1 of the BioCro II paper
[@10.1093/insilicoplants/diac003].  A few corrections to and
clarifications of the exposition given in that paper are given in the
comments below.

## The Code

```{r, original-example, fig.show="hold", out.width="50%"}
library(BioCro)
library(lattice)
library(knitr) # for kable(), which yields nicer-looking tables

## customized version of kable with defined options:
columns_to_print <- c('time', 'Q', 'mass_gain', 'Root', 'Leaf')
show_row_number <- FALSE
format <- list(scientific = FALSE, digits = 7)
cable <- function(x, ...) {
   kable(x[, columns_to_print],
         row.names = show_row_number,
         format.args = format,
         ...)
}

## Set plotting character globally
trellis.par.set("plot.symbol", list(pch = '.'))

################################################################################

parameters <- list(
    alpha_rue = 0.07, # kg / mol
    SLA = 25,         # m^2 / kg
    C_conv = 0.03,    # kg / mol
    f_leaf = 0.2,     # kg / kg
    f_root = 0.8,     # kg / kg
    timestep = 1      # s
)

initial_values <- list(
    Leaf = 1, # kg
    Root = 1  # kg
)

Q <- function(time) sin(time/3600/12 * pi) * 2000e-6 # mol / m^2 / s

times <- 0:(3600 * 12) # seconds

light_intensity <- data.frame(
    time = times,
    Q = Q(times)
)

result <- run_biocro(
    initial_values,
    parameters,
    light_intensity,
    'BioCro:example_model_mass_gain',
    'BioCro:example_model_partitioning'
)

xyplot(Q ~ time, data = result, xlab = "time (s)", ylab = "Q (mol / m^2 / s)")

xyplot(Leaf ~ time, data = result, xlab = "time (s)", ylab = "Leaf (kg)")
```

### A few comments

#### The formula for the driver

A somewhat plausible mechanistic model of the photosynthetic photon
flux density Q due to the sun would give it as being roughly
proportional to the sine of the angle the sun makes with the horizon.
Assuming 12 hours of daylight, with the sun passing directly overhead
(a condition that approximately holds at the equator during an
equinox), this angle (measured in radians) is approximately
`time/3600/12 * pi`, where `time` is the number of seconds that have
elapsed since sunrise.  (Here, an angle greater than `pi/2` is taken
to be the angle of the position of the sun measured from its position
at sunrise; this value thus takes a maximum value of `pi` at sunset.)
Assuming a maxumum flux density of `2000e-6 mol / m^2 /s` (attained at
solar noon when the sine is 1) then yields the function given for `Q`.

Note that instead of supplying `Q` as a driver, we could have instead
written a direct module that computed Q from the time of day.  The
latter would then be, for this example, the only driver.

#### The `timestep` parameter

Note that the timestep value given in the parameters is in seconds.
**But also note that almost all differential modules in the BioCro
library assume that timestep values are given in hours!**

If a timestep value is given in seconds rather than in hours, three
important things must be kept in mind:

1. _All_ differential modules used in the simulation must assume that
the timestep value is given in seconds.  Another way of saying this is
that the outputs of the differential modules used should represent the
rate of change of the output quantities *per second*.

2. In almost all cases, the output columns "doy", "hour", and possibly
"time" won't make sense and (if not filtered out) should be ignored.

3. In almost all cases, no module having `time` as input should be
used.  (See below for further discussion.)

The same precepts apply, _mutatis mutandi_, to any other timestep
units one may wish to use.

As always, the difference in time between any row _n_ in the _drivers_
dataframe and the row _n + 1_ that immediately follows it should be
equal to "timestep" (in whatever units _timestep_ has been given). BioCro
will check this is the case, and raise an error if not.


## First modification: 24 hour simulation

Although in the paper it is stated that "Light intensity is given as a
table of values at every second in a 24-h period,", the period here is
actually only 12 hours (time ranges from 0 to 3600 × 12 seconds = 12
hours).  As noted above, for the purposes of this example, it may be
assumed that time ranges from sunrise to sunset of a 12-hour day so
that "time" represents the number of seconds that have elapsed since
sunrise.

Alternatively, we could have assumed time to range over a 24-hour
period starting and ending at solar midnight, with Q being zero before
sunrise and after sunset.  This entails a slight modification to the
defining equations:

    times <- 0:(3600 * 24) # seconds

    Q <- function(time) pmax(0, sin((time/3600 - 6)/12 * pi) * 2000e-6) # mol / m^2 / s

(The "− 6" here adjusts the phase of the sine function so that the
positive values correspond to the daylight hours, assumed to begin at
6 a.m. solar time.)

Here is how the 24-hour version of the example would look.  A few rows
of the result table are printed at significant portions of the day: at
the beginning, just after midnight, when it is dark and there is no
mass accumulation; at dawn, when the rate of mass accumulation first
becomes non-zero; at noon, when the flux Q is highest; at dusk, when
the rate of mass accumulation again drops to zero; and at the end of
the simulation, when the final leaf and root mass are known.

As noted above, the values in the result for `doy` and `hour` are
nonsensical in the context of this example and so we have filtered
them out (along with `ncalls`, which is not of particular interest).
And the values in the `time` column should be interpreted as the
number of elapsed seconds from solar midnight (not the number of
(fractional) days that have elapsed since midnight at the start of
December 31, which is the correct interpretation in a "normal" BioCro
simulation).

```{r, 24-hour-version}
times <- 0:(3600 * 24) # seconds

Q <- function(time) {
    pmax(0, sin((time/3600 - 6)/12 * pi) * 2000e-6) # mol / m^2 / s
}

light_intensity <- data.frame(
    time = times,
    Q = Q(times)
)

result <- run_biocro(
    initial_values,
    parameters,
    light_intensity,
    'BioCro:example_model_mass_gain',
    'BioCro:example_model_partitioning'
)

## cable = customized version of kable (see above)
## kable = (nicely-formatted) knitr table
cable(result[1:4,]) # beginning
cable(result[seq(21591, 21661, 10),]) # around dawn
cable(result[43199:43203,]) # around mid day
cable(result[seq(64741, 64811, 10),]) # around dusk
cable(result[86398:86401,]) # end
```

In the following graphs, we have changed the labeling on the x axis so
that time is shown in hours rather than seconds.

```{r, plots-for-24-hour-version, fig.show="hold", out.width="50%"}
xyplot(Q ~ time / 3600, data = result, xlim = c(0, 24), xlab = "time (in hours)", ylab = "Q (mol / m^2 / s)")

xyplot(Leaf ~ time / 3600, data = result, xlim = c(0, 24), xlab = "time (in hours)", ylab = "Leaf (kg)")
```

## Second modification: Changing the time specification from seconds to days

The usual assumption in BioCro is that the `time` quantity represents
the time, in hours, that has elapsed from the beginning of the year,
not the number of seconds that have elapsed from the start of the simulation.

It follows, then, that the way we have been using the `time` variable
in the previous examples in inconsistent with the usual interpretation
of time in two respects: First, the units were in seconds rather than
days. Second, the time represented the amount of time elapsed from
the beginning of the simulation rather than an "absolute" time, that
is, a time value specifying a specific time of day on a specific day
of the year (a date and time on the calendar).
(There is no great harm in using the `time` variable in
this unconventional way of the previous examples, as long as we know
what we are doing and as long as we aren't using any modules that
assume the _conventional_ interpretation of the `time` variable.  In
fact, in simulations such as these ones, the result table is arguably
more readable when we have our time value in seconds or hours rather
than in days.)

The first of these inconsistencies is easily remedied by rescaling the
time values so that instead of running from 0 to 86400 (seconds), they
run from from 0 to 1 (days): we could simply divide each former time value
by 86400.  Now, the value of `time` would increase by 1/86400 (of a day!)
from row to row (an amount still corresponding to one second). However,
`run_biocro` requires `time` to be spaced by `timestep`.

```{r, consistent-durations, fig.show="hold", out.width="50%"}
light_intensity <- data.frame(
    time = times,
    Q = Q(times)
)

result <- run_biocro(
    initial_values,
    parameters,
    light_intensity,
    'BioCro:example_model_mass_gain',
    'BioCro:example_model_partitioning'
)

result$time <- result$time /86400

cable(result[1:4,]) # beginning
cable(result[seq(21591, 21661, 10),]) # around dawn
cable(result[43199:43203,]) # around mid day
cable(result[seq(64741, 64811, 10),]) # around dusk
cable(result[86398:86401,]) # end

xyplot(Q ~  24 * time , data = result, xlab = "hour", ylab = "Q (mol / m^2 / s)")

xyplot(Leaf ~ 24 * time , data = result, xlab = "hour", ylab = "Leaf (kg)")
```

## Third modification: Making the `time` variable consistent with a reasonable time of year

The simulation just shown takes care of making the time step
consistent with the time duration represented by the time column in
its usual interpretation.  But the `time` value of 0 corresponds (in
the conventional BioCro interpretation) to midnight January 1st, and
this is not consistent with a sun passing directly
overhead during the course of a 12 hour day, even at an equatorial
location.  We may remedy this by shifting the time values forward by
79 days to the time of the March equinox.  Direct modules that rely on this
meaning for time, such as `solar_position_michalsky`, could now be
used if we wished (`time` represents a date and time).
(But we still can not mix in derivative modules,
such as the thermal time modules, that use `time` as input, since
these, unlike the `example_model_partitioning` module we are using,
assume the timestep value to be in hours, not seconds.)

```{r, realistic-time-of-year, fig.show="hold", out.width="50%"}
light_intensity <- data.frame(
    time = times +  79 * 86400,
    Q = Q(times)
)

result <- run_biocro(
    initial_values,
    parameters,
    light_intensity,
    'BioCro:example_model_mass_gain',
    'BioCro:example_model_partitioning'
)


cable(result[1:4,]) # beginning
cable(result[seq(21591, 21661, 10),]) # around dawn
cable(result[43199:43203,]) # around mid day
cable(result[seq(64741, 64811, 10),]) # around dusk
cable(result[86398:86401,]) # end

xyplot(Q ~ time/3600 -  24 * (79 - 1), data = result, xlim = c(0, 24), xlab = "hour of the 79th day of the year\n(March 20)", ylab = "Q (mol / m^2 / s)")

xyplot(Leaf ~ time/3600 - 24 * (79 - 1) , data = result, xlim = c(0, 24), xlab = "hour of the 79th day of the year\n(March 20)", ylab = "Leaf (kg)")
```

## References
