---
title: "Energy Balance, Transpiration, and Leaf Temperature"
output:
  bookdown::html_vignette2:
    toc: true
    number_sections: true
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Energy Balance, Transpiration, and Leaf Temperature}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{bookdown}
bibliography: references/energy_balance_references.bib
csl: chicago-author-date.csl
link-citations: yes
pkgdown:
  as_is: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.width = 5,
  fig.height = 5,
  fig.align = "center"
)
```

# Canopy Energy Balance {#generalEB}

Energy balance is discussed in many sources, such as @thornley_plant_1990 and
@campbell_introduction_1998, the two main references for BioCro models. Here we
begin with a general discussion, mostly using the notation from Chapter 14 of
@thornley_plant_1990, but also borrowing a bit from Chapters 12 and 14 of
@campbell_introduction_1998.

First, the total light energy (shortwave and longwave) absorbed by the canopy
($J_a$) is all dissipated or stored by the following processes:

 * Long-wave energy loss to the environment

 * Convective heat transfer between the canopy and the air

 * Evaporation of water from the interior surfaces of the leaves

 * Conductive heat transfer between the canopy and the ground

 * Canopy metabolism

 * Heat storage by the canopy

Mathematically, we can express this idea as

\begin{equation}
    J_a - R_L - H - \lambda E - G - M = q, (\#eq:EB1)
\end{equation}

where $R_L$ is the rate of long-wave energy loss, $H$ is the rate of conductive
heat transfer, $\lambda$ is the latent heat of vaporization of water, $E$ is the
transpiration rate, $G$ is the rate of heat loss to the ground, $M$ is the rate
of energy used by canopy metabolism, and $q$ is the rate of heat storage by the
canopy. Most of the quantities above are energy flux densities with typical
units of J m$^{-2}$ s$^{-1}$. The only exception is $\lambda$, which typically
has units of J kg$^{-1}$. Equation \@ref(eq:EB1) is a crop-specific version of
the more general principle of
[conservation of energy](https://en.wikipedia.org/wiki/Conservation_of_energy),
and it already begins to highlight the connection between energy and
transpiration.

The term "energy balance" is synonymous with "steady-state energy flow." Thus,
energy balance calculations are always made under the assumption that $q = 0$.
Another simplification is that the metabolic processes of photosynthesis and
respiration only dissipate a small fraction of the incident energy, and can thus
be ignored. Finally, we can combine $J_a$, $R_L$, and $G$ into a single term
($\phi_N$):

\begin{equation}
    \phi_N = J_a - R_L - G. (\#eq:phiN)
\end{equation}

With these changes, \@ref(eq:EB1) becomes

\begin{equation}
    \phi_N - H - \lambda E = 0. (\#eq:EB2)
\end{equation}

In other words, $\phi_N$ is the net energy available to the crop for driving
transpiration and sensible heat loss.

The terms $H$ and $E$ are related to the conductance of heat and water vapor
from the canopy to the ambient environment. In fact, they can be both be
expressed using one-dimensional diffusion equations:

\begin{equation}
    H = \rho c_p \left( T_l - T_a \right) g_H (\#eq:H)
\end{equation}

and

\begin{equation}
    E = \left(\rho_{vl} - \rho_{va} \right) g_W, (\#eq:E)
\end{equation}

where $\rho$ is the density of dry air (kg m$^{-3}$), $c_p$ is the specific heat
capacity of dry air (J kg$^{-1}$ K$^{-1}$), $T_a$ and $T_l$ are the temperatures
of the ambient air and the leaves in canopy, respectively (K), $\rho_{va}$ and
$\rho_{vl}$ are the water vapor densities in the ambient air and within the
leaves, respectively (kg water m$^{-3}$), and $g_H$ and $g_W$ are the
conductances to heat and water vapor (m s$^{-1}$). Note that temperatures are
expressed in K here to ensure that the product $\rho c_p T$ is an energy density
with units J m$^{-3}$ (even though the temperature difference $T_l - T_a$ is
numerically identical whether K or $^{\deg}$C are used for the temperatures).
Thus, both Equation \@ref(eq:H) and \@ref(eq:E) express a flux density as the
product of a conductance and a density gradient.

For a healthy leaf with sufficient access to water, in an environment with
reasonably low vapor pressure deficit, the air in its intercellular spaces is
close to full saturation [@wong_humidity_2022]. Thus, here we will make a
simplifying assumption that $\rho_{vl} = \rho_{sat}(T_l)$, where
$\rho_{sat}(T_l)$ is the saturation water vapor density at the leaf temperature.
With this assumption, Equation \@ref(eq:E) can be expressed as

\begin{equation}
    E = \left[ \rho_{sat}(T_l) - \rho_{va} \right] g_W. (\#eq:E2)
\end{equation}

Combining Equations \@ref(eq:EB2), \@ref(eq:H), and \@ref(eq:E2), we can now
express the energy balance requirement as

\begin{equation}
    \phi_N - \rho c_p \left( T_l - T_a \right) g_H -
        \lambda \left[ \rho_{sat}(T_l) - \rho_{va} \right] g_W = 0. (\#eq:EB3)
\end{equation}

This is a powerful equation because all of its terms either (i) can be taken as
constant for a given set of environmental conditions or (ii) have a known
dependence on the leaf temperature. Because of this, we can consider Equation
\@ref(eq:EB3) as a way to calculate the temperature of leaves in the canopy.
Once $T_l$ is known, Equation \@ref(eq:E2) can be used to calculate the
transpiration rate $E$. Often this is the main goal of an energy balance
calculation, since the transpiration rate determines crop water use.

However, Equation \@ref(eq:EB3) cannot be solved analytically as it is written.
One main option is to solve it using a numerical approach, while another is to
approximate some of its terms in a way that enables an analytical expression for
$T_l$ and $E$. The following sections will discuss different approaches to this
problem with varying levels of complexity.

# A Penman-Monteith Equation {#pmthornley}

Here we continue following Chapter 12 of @thornley_plant_1990, which
derives analytical expressions for $T_l$ and $E$ based on energy balance
(Equation \@ref(eq:EB3)). In this approach, we make a few more assumptions:

 * We treat the diffusion path for water vapor as consisting of two sequential
   components: (i) diffusion from the intercellular spaces across the stomata to
   the leaf surfaces (with conductance $g_c$) and (ii) diffusion from the leaf
   surfaces across the boundary layer to the bulk atmosphere (with conductance
   $g_a$).

 * We assume that the flows of water vapor and heat across the boundary
   layer are driven by the same mechanical processes.

 * We assume that $\phi_N$, $g_a$, and $g_c$ do not depend on $T_l$.

 * We assume that the temperature difference between the ambient air and the
   leaves in the canopy ($\Delta T = T_l - T_a$) is small.

When $\Delta T$ is small, it is possible to linearize the equation for
saturation water vapor density $\rho_{sat}(T)$. This is an approach first taken
by Penman. To do this, we first express $\rho_{sat}(T_l)$ as a
[Taylor series](https://en.wikipedia.org/wiki/Taylor_series) centered on $T_a$:

\begin{equation}
    \rho_{sat}(T_l) = \rho_{sat}(T_a) +
        \frac{d \rho_{sat}}{d T} \bigg\rvert_{T_a} \cdot \Delta T +
        \frac{1}{2!} \cdot \frac{d^2 \rho_{sat}}{d T^2} \bigg\rvert_{T_a} \cdot \left( \Delta T \right)^2 + ...
\end{equation}

If $\Delta T$ is small, then $(\Delta T)^2$ is even smaller, and we can consider
it and any higher-order terms to be negligable. In this case, we can keep just
the first two terms of the Taylor series:

\begin{equation}
    \rho_{sat}(T_l) \approx \rho_{sat}(T_a) + s \Delta T (\#eq:linearRho),
\end{equation}

where $s = \frac{d \rho_{sat}}{d T} \bigg\rvert_{T_a}$ is the slope of
$\rho_{sat}(T)$ at $T = T_a$. With this simplification, Equation \@ref(eq:E2)
becomes

\begin{equation}
    E = \left[ \rho_{sat}(T_a) + s \left( \Delta T \right) - \rho_{va} \right] g_W =
        \left( \Delta \rho_{va} + s \Delta T \right) g_W (\#eq:E3)
\end{equation}

where $\Delta \rho_{va} = \rho_{sat}(T_a) - \rho_{va}$ is the vapor density
deficit of the ambient air.

Because the stomata and boundary layer are in series, the total conductance to
water vapor ($g_W$) is related to $g_a$ and $g_c$ according to

\begin{equation}
    \frac{1}{g_W} = \frac{1}{g_a} + \frac{1}{g_c}, (\#eq:gW)
\end{equation}

Additionally, because the same mechanisms of conduction and convection drive the
flow of both water vapor and heat in the boundary layer, these conductances are
identical:

\begin{equation}
    g_H = g_a. (\#eq:gH)
\end{equation}

Combining Equations \@ref(eq:EB3), \@ref(eq:linearRho), \@ref(eq:gW), and
\@ref(eq:gH), we can arrive at an expression of conservation of energy for a
crop canopy where only water vapor conductances are used:

\begin{equation}
    \phi_N - \rho c_p \Delta T g_a -
        \frac{\lambda \left( \Delta \rho_{va} + s \Delta T \right)}{1 / g_a + 1 / g_c} = 0. (\#eq:EB4)
\end{equation}

These may not look simple. Nevertheless, with some algebra, Equation
\@ref(eq:EB4) can be solved for the temperature difference $\Delta T$:

\begin{equation}
    \Delta T = \frac{\phi_N \left( 1 / g_a + 1 / g_c \right) - \lambda \Delta \rho_{va}}{\lambda \left[ s + \gamma \left( 1 + g_a / g_c \right) \right]} (\#eq:DT1)
\end{equation}

where $\gamma = \frac{\rho c_p}{\lambda}$ is the "psychrometric parameter." This
is Equation 14.11e in @thornley_plant_1990. Note that by assumption, everything
on the right-hand side is independent of the leaf temperature. Values of $s$,
$\rho$, $c_p$, etc can be found in reference tables, so if values of $\phi_N$,
$g_a$, etc, are known or can be estimated, then Equation \@ref(eq:DT1) can be
used to calculate $\Delta T$ (and hence $T_l$). It can also be plugged in to
Equation \@ref(eq:E3) to calculate the canopy transpiration rate $E$:

\begin{equation}
    E = \frac{s \phi_N + \lambda \gamma g_a \Delta \rho_{va}}{\lambda \left[ s + \gamma \left(1 + g_a / g_c \right) \right]}, (\#eq:PM1)
\end{equation}

This is Equation 14.4k from @thornley_plant_1990, where it is described as "the
Penman-Monteith equation." However, just one page later, @thornley_plant_1990
gives an alternate version (Equation 14.5d). As it turns out, there are many
different equations referred to as "the Penman-Monteith equation" throughout
textbooks and the plant physiology literature. Thus, we prefer to call it "a
Penman-Monteith equation" to indicate it is just one example from a group of
potential Penman-Monteith equations. In the next section, we will discuss
another version.

# Another Penman-Monteith Equation {#pmcampbell}

The assumptions given at the start of the [previous section](#pmthornley) are
not the only way to build analytical expressions for $\Delta T$ and $E$ from
Equation \@ref(eq:EB3). In fact, a different approach can be found in Chapter 14
of @campbell_introduction_1998. Here we make some alternate assumptions:

 * We assume that $J_a$ and $G$ do not depend on leaf temperature, but we use
   the [Stefan-Boltzmann law](https://en.wikipedia.org/wiki/Stefan%E2%80%93Boltzmann_law)
   to calculate $R_L$ from $T_l$.

 * We assume that $g_H$ and $g_W$ do not depend on leaf temperature.

 * We assume that the temperature difference between the ambient air and the
   leaves in the canopy ($\Delta T = T_l - T_a$) is small.

Applying the Stefan-Boltzmann law, we now have an expression for $R_L$:

\begin{equation}
    R_L = \epsilon_s \sigma T_l^4,
\end{equation}

where $\epsilon_s$ is the canopy surface emissivity and $\sigma$ is the
Stefan-Boltzmann constant.

Because $\Delta T$ is small, we can linearize the saturation water vapor density
as before (Equation \@ref(eq:linearRho)). We can also linearize $R_L$ by noting
that

\begin{equation}
    T_L^4 = (T_a + \Delta T)^4 = T_a^4 + 4 T_a^3 \Delta T +
        6 T_a^2 (\Delta T)^2 + 4 T_a (\Delta T)^3 + (\Delta T)^4.
\end{equation}

We can again neglect terms with $\Delta T^2$ or higher powers, yielding

\begin{equation}
    R_L \approx \epsilon_s \sigma \left( T_a^4 + 4 T_a^3 \Delta T \right) =
        \epsilon_s \sigma T_a^4 - \rho c_p \Delta T g_r, (\#eq:linearRL)
\end{equation}

where $g_r = 4 \epsilon_s \sigma T_a^3 / ( \rho c_p)$ is a "radiative
conductance" with units m s$^{-1}$. Substituing this linearization into Equation
\@ref(eq:phiN) yields

\begin{equation}
    \phi_N = J_a - \epsilon_s \sigma T_a^4 - \rho c_p \Delta T g_r  - G, (\#eq:newphilinear)
\end{equation}

Combining Equations \@ref(eq:EB3), \@ref(eq:linearRho), and
\@ref(eq:newphilinear), we have yet another a new version of the canopy energy
balance equation:

\begin{equation}
    J_a - \epsilon_s \sigma T_a^4 - \rho c_p \Delta T g_r  - G - \rho c_p \Delta T g_H -
        \lambda \left( \Delta \rho_{va} + s \Delta T \right) g_W = 0.
\end{equation}

This can also be written as

\begin{equation}
    J_a - \epsilon_s \sigma T_a^4 - G - \rho c_p g_{Hr} \Delta T -
        \lambda \left( \Delta \rho_{va} + s \Delta T \right) g_W = 0, (\#eq:EB5)
\end{equation}

where $g_{Hr} = g_r + g_H$ is a combined conductance for radiation and heat.
This expression of energy balance can now be solved for $\Delta T$ to produce an
equation analgous to Equation \@ref(eq:DT1):

\begin{equation}
    \Delta T = \frac{\left( J_a - \epsilon_s \sigma T_a^4  - G \right) / g_W - \lambda \Delta \rho_{va}}{\lambda \left[ s + \gamma (g_{Hr} / g_W) \right]}. (\#eq:DT2)
\end{equation}

Equation \@ref(eq:DT2) can be substituted into Equation \@ref(eq:E3) to
calculate the canopy transpiration rate $E$:

\begin{equation}
    E = \frac{s \left( J_a - \epsilon_s \sigma T_a^4  - G \right) + \lambda \gamma g_{Hr} \Delta \rho_{va}}{\lambda \left[ s + \gamma (g_{Hr} / g_W) \right]}. (\#eq:PM2)
\end{equation}

This equation is analagous to, but different from, Equation \@ref(eq:PM1).
Disregarding differences in notation, it is also identical to Equation 14.12
from @campbell_introduction_1998, where it is described as follows:

> This is the well known and widely used Penman-Monteith equation for estimating
> evapotranspiration from plant communities.

Yet, it is substantially different from the Penman-Monteith equation given in
@thornley_plant_1990 due to the different assumptions and linearizations made
while deriving it.

# The Penman-Monteith Approach

Sections \@ref(pmthornley) and \@ref(pmcampbell) illustrate two possible ways to
make approximations that enable Equations \@ref(eq:EB3) and \@ref(eq:E2) to be
solved for $T_l$ and $E$. However, there are many possible choices that could,
in principle, each lead to slightly different versions of the equations:

 * Energy balance can be considered at the canopy or leaf scale.

 * Water vapor concentrations can be expressed as vapor densities, partial
   pressures, or mole fractions.

 * $R_L$ can be linearized or considered as a leaf-temperature-independent
   component of $\phi_N$.

 * The diffusion path for water vapor can be separated into distinct sequential
   steps or treated as a single process.

This list is not necessarily exhaustive, but serves to illustrate the idea that
there is not one single Penman-Monteith equation. However, each potential
version of the equation includes a linearization of $p_{sat}(T_l)$ and possibly
$R_L(T_l)$, approximating each of these as a linear function of $\Delta T$.
This linearization is what enables the energy balance equation to be solved for
$T_l$. Thus, this "Penman-Monteith approach" is common to all of these equations
even when there are differences between the equations themselves.

# Energy Balance in BioCro

In Sections \@ref(generalEB), \@ref(pmthornley), and \@ref(pmcampbell), the
canopy transpiration rate $E$ depends on canopy-level conductances: $g_H$,
$g_W$, $g_a$, $g_c$, and $g_r$. In practice, these conductances are difficult
to estimate _a priori_ because they depend on gradients of wind speed, light
intensity, and other factors within a plant canopy. In BioCro, we take an
alternate approach where transpiration rates are estimated at the leaf level
and then scaled up to the canopy level. This enables us to use models for
leaf-level conductance (such as the Ball-Berry model for stomatal conductance)
and also to account for variations in key inputs like $J_a$ throughout the
canopy. In this framework, the total canopy transpiration rate is given by

\begin{equation}
    E_{canopy} = \int_{\ell = 0}^{L} \left( E_{leaf}^{sunlit}(\ell) \cdot f_{sunlit}(\ell) +
        E_{leaf}^{shaded}(\ell) \cdot f_{shaded}(\ell) \right) d \ell, (\#eq:ECanopyInt)
\end{equation}

where $E_{leaf}^{sunlit}(\ell)$ and $E_{leaf}^{shaded}(\ell)$ are the
leaf-level transpiration rates per leaf area for sunlit and shaded
leaves at cumulative leaf area index (cLAI) $\ell$, $f_{sunlit}(\ell)$ and
$f_{shaded}(\ell)$ are the fractions of sunlit and shaded leaf area at $\ell$,
and $L$ is the total canopy leaf area index (LAI).

The equations derived above using the Penman-Monteith approach assume that $g_W$
is independent of both leaf temperature and incident light intensity.
Realistically, this assumption is likely to be violated because stomatal
conductance changes dynamically with photosynthesis, which in turn changes
dynamically with leaf temperature and incident light intensity. There is also a
buoyancy component to boundary layer conductance that depends on leaf
temperature. The equations used to calculate these conductances are complex and
not readily amenable to the types of linearizations employed in the
Penman-Monteith approach. So, in BioCro, we take a hybrid approach to energy
balance where $T_l$ is varied iteratively to simultaneously determine $g_W$,
$T_l$, and $E$ using Equation \@ref(eq:DT1). See the `leaf_energy_balance()`
function in BioCro's `C++` code for more details.

Because $E_{leaf}$ is calculated numerically in BioCro, it is not possible to
obtain an analytical expression for $E_{canopy}$ using Equation
\@ref(eq:ECanopyInt). Instead, the integral is approximated using a
[Riemann sum](https://en.wikipedia.org/wiki/Riemann_sum):

\begin{equation}
    E_{canopy} = \sum_{i = 0}^{N-1} \left( E_{leaf}^{sunlit}(\ell_i) \cdot f_{sunlit}(\ell_i) +
        E_{leaf}^{shaded}(\ell_i) \cdot f_{shaded}(\ell_i) \right) \cdot \frac{L}{N}. (\#eq:ECanopySum)
\end{equation}

Here the cLAI interval $(0, L)$ has been divided into $N$ sub-intervals of
equal width ($L / N$) and $\ell_i = (i + 0.5) \cdot L / N$ is the midpoint of
the $i^{th}$ sub-interval.

Values of $T_l$ and $E$ calculated from energy balance are sensitive to the
conductances across the stomata and boundary layer. Yet, the boundary layer
conductance in particular is difficult to estimate because it depends on details
such as wind speed, leaf shape, and other factors that change dynamically in
realistic conditions and cannot be easily quantified. In BioCro, we consider the
flow of water vapor from the leaf's intercellular spaces to the bulk air in the
atmosphere to consist of three sequential steps as the gas crosses the stomata,
the leaf boundary layer, and finally the canopy boundary layer. The latter two
conductances combine to form the overall boundary layer conductance:

\begin{equation}
    \frac{1}{g_b} = \frac{1}{g_{b,leaf}} + \frac{1}{g_{b,canopy}},
\end{equation}

where $g_{b,leaf}$ and $g_{g,canopy}$ are the conductances across the leaf and
canopy boundary layers.

$g_{b,leaf}$ is calculated using a model described in @nikolov_coupling_1995. It
treats the leaves in the canopy as individual flat objects in a moving air
stream, and considers free and forced convection, where free convection is
driven by wind and forced convection is driven by temperature gradients. In
BioCro, wind speed within the canopy is modeled as an exponential decay with the
cumulative leaf area index, and this depth-dependent wind speed is used to
calculate $g_{b,leaf}$. Because of this, and a dependence on stomatal
conductance, $g_{b,leaf}$ changes throughout the canopy and is different for
sunlit and shaded leaves.

On the other hand, $g_{b,canopy}$ is calculated using a model described in
@thornley_plant_1990. It treats the canopy as an infinitely wide rough surface
covering the ground, and only considers turbulent flow, so $g_{b,canopy}$
primarily depends on the wind speed measured above the canopy. The model
equations predict $g_{b,canopy} = 0$ when wind speed is zero. In reality,
temperature-driven convection would occur at low wind speeds, setting a lower
limit for $g_{b,canopy}$ as in the model for $g_{b,leaf}$. To account for this
in a simple way, a minimum value of $g_{b,canopy}$ is enforced in BioCro. The
minimum conductance is a parameter that can be varied by the user.
$g_{b,canopy}$ is the same for sunlit and shaded leaves at all depths within the
canopy.

The numeric technique used to solve for $T_l$ in BioCro (fixed-point iteration)
does not always converge to a reasonable solution. To prevent extremely high or
low leaf temperatures, $\Delta T$ is restricted to lie between -30$^{\circ}$C
and 30$^{\circ}$C. This restriction tends to be enforced only in unusual
situations where stomatal conductance is very low (and hence assimilation is low
as well), and thus has little impact on the overall result of a simulation.

## BioCro Examples

```{r}
library(BioCro)
library(lattice)
```

BioCro's leaf-level energy balance model:
```{r}
transpiration_res <- module_response_curve(
    'BioCro:leaf_evapotranspiration',
    list(
        absorbed_shortwave = 250,
        atmospheric_pressure = 101325,
        temp = 25,
        canopy_height = 1,
        leafwidth = 0.1,
        min_gbw_canopy = 0.005,
        rh = 0.6,
        wind_speed_height = 10
    ),
    expand.grid(
        windspeed = seq(0, 2, by = 0.005),
        Gs = c(0.05, 0.1, 0.2, 0.4, 0.8)
    )
)

xyplot(
    TransR ~ windspeed,
    group = Gs,
    data = transpiration_res,
    type = 'l',
    auto = TRUE,
    xlab = 'Wind speed (m / s)',
    ylab = 'Transpiration rate (mmol / m^2 / s)',
    ylim = c(0, 5),
    main = 'Transpiration rate for different\nstomatal conductances (mol / m^2 / s)'
)

xyplot(
    leaf_temperature ~ windspeed,
    group = Gs,
    data = transpiration_res,
    type = 'l',
    auto = TRUE,
    xlab = 'Wind speed (m / s)',
    ylab = 'Leaf temperature (degrees C)',
    ylim = c(0, 100),
    main = 'Leaf temperature for different\nstomatal conductances (mol / m^2 / s)'
)
```

BioCro's model for canopy boundary layer conductance:
```{r}
canopy_gbw_res <- module_response_curve(
    'BioCro:canopy_gbw_thornley',
    list(
        min_gbw_canopy = 0,
        windspeed_height = 10
    ),
    expand.grid(
        windspeed = seq(0, 2, by = 0.025),
        canopy_height = seq(0.5, 2, by = 0.5)
    )
)

xyplot(
    gbw_canopy ~ windspeed,
    group = canopy_height,
    data = canopy_gbw_res,
    type = 'l',
    auto = TRUE,
    xlab = 'Wind speed (m / s)',
    ylab = 'Canopy boundary layer conductance (m / s)',
    ylim = c(0, 0.02),
    main = 'Canopy gbw for different canopy heights'
)
```

BioCro's model for leaf boundary layer conductance:
```{r}
leaf_gbw_res <- module_response_curve(
    'BioCro:leaf_gbw_nikolov',
    list(
        air_pressure = 101325,
        leafwidth = 0.1,
        temp = 25,
        leaf_temperature = 26,
        rh = 0.6
    ),
    expand.grid(
        windspeed = seq(0, 1, by = 0.005),
        Gs = c(0.05, 0.1, 0.2, 0.4, 0.8)
    )
)

xyplot(
    gbw_leaf ~ windspeed,
    group = Gs,
    data = leaf_gbw_res,
    type = 'l',
    auto = TRUE,
    xlab = 'Wind speed (m / s)',
    ylab = 'Leaf boundary layer conductance (m / s)',
    ylim = c(0, 0.02),
    main = 'Leaf gbw for different stomatal conductances\n(Gs; mol / m^2 / s)'
)
```

# References
