---
title: "BioCro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BioCro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  error = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 5,
  fig.align = "center"
)
```

## Overview

BioCro is a model that predicts plant growth over time given crop-specific
parameters and environmental data as input.

It uses models of key physiological and biophysical processes underlying plant
growth, and has previously been used for predicting biomass yield and leaf area
index of switchgrass and miscanthus. In 2022, BioCro was reorganized to take a
truly modular approach to modeling and a new soybean model was developed.

## An example

The `run_biocro()` function accepts initial values, parameters, climate
variables, and sets of modules to run. It returns the results in a data frame.

```{r}
library(BioCro)
library(lattice)
weather_data <- add_time_to_weather_data(soybean_weather$'2002') # required 

result <- with(soybean, {run_biocro(
  initial_values,
  parameters,
  weather_data,
  direct_modules,
  differential_modules,
  ode_solver
)})

xyplot(Stem + Leaf ~ TTc, data = result, type='l', auto = TRUE)
```

There are parameters and modules for soybean (_Glycine max_), miscanthus
(_Miscanthus_ x _giganteus_), and willow (_Saliceae salix_).

## Getting Started

The best place to get started with BioCro is the
[public BioCro Documentation web site](https://biocro.github.io), which includes
the following:

- Installation instructions

- A "Practical Guide to BioCro" and other articles in the "Articles" dropdown
  menu

- Documentation of the C++ code, including notes on the biological models used
  in BioCro, in the "C++ Library" dropdown menu

- Documentation for BioCro developers and maintainers in the "Developer's
  Manual" dropdown menu

There is also a separate [page that documents all of the quantities
used by the Standard BioCro Module
Library](https://biocro.github.io/BioCro-documentation/quantity_docs/quantities.html).

## License

The `BioCro` R package is licensed under the MIT license, while the BioCro C++
framework is licensed under version 3 or greater of the GNU Lesser General
Public License (LGPL). This scheme allows people to freely develop models for
any use (public or private) under the MIT license, but any changes to the
framework that assembles and solves models must make source code changes
available to all users under the LGPL. See `LICENSE.note` for more details.
