---
title: "BioCro II Paper: Section 1.1 Example"
output:
  html_document:
    toc: true

vignette: >
  %\VignetteIndexEntry{BioCro II Paper: Section 1.1 Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

nocite: "[@*]"
references:
- author:
  - family: Lochocki
    given: Edward B
  - family: Rohde
    given: Scott
  - family: Jaiswal
    given: Deepak
  - family: Matthews
    given: Megan L
  - family: Miguez
    given: Fernando
  - family: Long
    given: Stephen P
  - family: McGrath
    given: Justin M
  container-title: in silico Plants
  doi: 10.1093/insilicoplants/diac003
  id: 10.1093/insilicoplants/diac003
  issn: 2517-5025
  issue: 1
  issued: 2022-02
  note: diac003
  title: "[BioCro II: a software package for modular crop growth
    simulations]{.nocase}"
  type: article-journal
  url: "https://doi.org/10.1093/insilicoplants/diac003"
  volume: 4
---
```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = TRUE,
  autodep = TRUE
)
```

## Introduction

This is a demonstration of the example discussed in Section 1.1 and
Appendix 1 of the seminal BioCro II paper
[@10.1093/insilicoplants/diac003].  A few corrections to and
clarifications of the exposition given in the paper are given in the
comments below.

## The Code

```{r, original-example, fig.show="hold", out.width="50%"}
library(BioCro)
library(lattice)
library(knitr) # for kable(), which yields nicer-looking tables

## Set plotting character globally
trellis.par.set("plot.symbol", list(pch = '.'))

################################################################################

parameters <- list(
    alpha_rue = 0.07, # kg / mol
    SLA = 25,         # m^2 / kg
    C_conv = 0.03,    # kg / mol
    f_leaf = 0.2,     # kg / kg
    f_root = 0.8,     # kg / kg
    timestep = 1      # s
)

initial_values <- list(
    Leaf = 1, # kg
    Root = 1  # kg
)

Q <- function(time) sin(time/3600/12 * pi) * 2000e-6 # mol / m^2 / s

times <- 0:(3600 * 12) # seconds

light_intensity <- data.frame(
    time = times,
    Q = Q(times)
)

result <- run_biocro(
    initial_values,
    parameters,
    light_intensity,
    'BioCro:example_model_mass_gain',
    'BioCro:example_model_partitioning'
)

xyplot(Q ~ time, data = result, xlab = "time (s)", ylab = "Q (mol / m^2 / s)")

xyplot(Leaf ~ time, data = result, xlab = "time (s)", ylab = "Leaf (kg)")
```

### A few comments

#### The formula for the driver

A somewhat plausible mechanistic model of the photosynthetic photon
flux density Q due to the sun would give it as being roughly
proportional to the sine of the angle the sun makes with the horizon.
Assuming 12 hours of daylight, with the sun passing directly overhead,
(a condition that approximately holds at the equator during an
equinox), this angle (measured in radians) is approximately
`time/3600/12 * pi`, where `time` is the number of seconds that have
elapsed since sunrise.  (Here, an angle greater than `pi/2` is taken
to be the angle of the position of the sun measured from its position
at sunrise; this value thus takes a maximum value of `pi` at sunset.)
Assuming a maxumum flux density of `2000e-6 mol / m^2 /s` (attained at
solar noon when the sine is 1) then yields the function given for `Q`.

Note that instead of supplying `Q` as a driver, we could have instead
written a direct module that computed Q from the time of day.  The
latter would then be, for this example, the only driver.

#### The `timestep` parameter

Note that the timestep value given in the parameters is in seconds.
**But also note that almost all differential modules in the BioCro
library assume that timestep values are given in hours!**

If a timestep value is given in seconds rather than in hours, three
important things must be kept in mind:

1. _All_ differential modules used in the simulation must assume that
the timestep value is given in seconds.  Another way of saying this is
that the outputs of the differential modules used should represent the
rate of change of the output quantities *per second*.

2. In almost all cases, the output columns "doy", "hour", and possibly
"time" won't make sense and (if not filtered out) should be ignored.

3. In almost all cases, no module having "doy", "hour", or "time" as
input should be used.  (See below for further discussion.)

The same precepts apply, _mutatis mutandi_, to any other timestep
units one may wish to use.

As always, the difference in time between any row n in the _drivers_
dataframe and the row n + 1 that immediately follows it should be
equal to "timestep" (in whatever units _timestep_ has been given).


## First modification: 24 hour simulation

Although in the paper it is stated that "Light intensity is given as a
table of values at every second in a 24-h period,", the period here is
actually only 12 hours (time ranges from 0 to `3600 * 12` seconds = 12
hours).  As noted above, for the purposes of this example, it may be
assumed that time ranges from sunrise to sunset of a 12-hour day so
that "time" represents the number of seconds that have elapsed since
sunrise.

Alternatively, we could have assumed time to range over a 24-hour
period starting and ending at solar midnight, with Q being zero before
sunrise and after sunset.  This entails a slight modification to the
defining equations:

    times <- 0:(3600 * 24) # seconds

    Q <- function(time) pmax(0, sin((time/3600 - 6)/12 * pi) * 2000e-6) # mol / m^2 / s

(The "- 6" here adjusts the phase of the sine function so that the
positive values correspond to the daylight hours, assumed to begin at
6 a.m. solar time.)

Here is how the 24-hour version of the example would look.  A few rows
of the result table are printed at significant portions of the day: at
the beginning, just after midnight, when it is dark and there is no
mass accumulation; at dawn, when the rate of mass accumulation first
becomes non-zero; at noon, when the flux Q is highest; at dusk, when
the rate of mass accumulation again drops to zero; and at the end of
the simulation, when the final Leaf and Root mass are known.

As noted above, **_the values in the result for `doy` and `hour`
should be ignored!  They are nonsensical in the context of this
example!_** And the values in the `time` column should be interpreted
as the number of elapsed seconds from solar midnight (not the number
of (fractional) days that have elapsed since midnight at the start of
December 31, which is the correct interpretation for a "normal" BioCro
simulation).

```{r, 24-hour-version}
times <- 0:(3600 * 24) # seconds

Q <- function(time) {
    pmax(0, sin((time/3600 - 6)/12 * pi) * 2000e-6) # mol / m^2 / s
}

light_intensity <- data.frame(
    time = times,
    Q = Q(times)
)

result <- run_biocro(
    initial_values,
    parameters,
    light_intensity,
    'BioCro:example_model_mass_gain',
    'BioCro:example_model_partitioning'
)

## kable = (nicely-formatted) knitr table
kable(result[1:4,]) # beginning
kable(result[seq(21590, 21660, 10),]) # around dawn
kable(result[43198:43202,]) # around mid day
kable(result[seq(64740, 64810, 10),]) # around dusk
kable(result[86398:86401,]) # end
```

In the following graphs, we have changed the labeling on the x axis so
that time is shown in hours rather than seconds.

```{r, plots-for-24-hour-version, fig.show="hold", out.width="50%"}
xyplot(Q ~ time / 3600, data = result, xlim = c(0, 24), xlab = "time (in hours)", ylab = "Q (mol / m^2 / s)")

xyplot(Leaf ~ time / 3600, data = result, xlim = c(0, 24), xlab = "time (in hours)", ylab = "Leaf (kg)")
```

## Second modification: Making the `doy` and `hour` duration consistent with the duration calculated from `timestep`

As noted above, in most cases the outputs "doy" and "hour" should be
ignored if timestep is not given in units of hours.  This is because
in most such cases, the difference in the values of doy and hour from
one row to the next will not correspond to a difference in time of one
timestep.  In the 24-hour simulation just shown, for example, 86400
one-second timesteps elapse between the beginning time of the
simulation and the ending time, correspond to a duration of 24 hours.
But the value of doy ranges from 0 at the beginning to 86400 at the
end, corresponding to a duration of as many days, and the value shown
for _hour_ is always zero.

We could, however, remedy this inconsistency by ensuring that the
values of time in our dataframe correspond to units of days but
that the value, nevertheless, increases by only an amount
corresponding to one second from row to row:

```{r, consistent-durations, fig.show="hold", out.width="50%"}
light_intensity <- data.frame(
    time = times/86400,
    Q = Q(times)
)

result <- run_biocro(
    initial_values,
    parameters,
    light_intensity,
    'BioCro:example_model_mass_gain',
    'BioCro:example_model_partitioning'
)

kable(result[1:4,]) # beginning
kable(result[seq(21590, 21660, 10),]) # around dawn
kable(result[43198:43202,]) # around mid day
kable(result[seq(64740, 64810, 10),]) # around dusk
kable(result[86398:86401,]) # end

xyplot(Q ~ (hour + 24 * doy), data = result, xlim = c(0, 24), xlab = "hour", ylab = "Q (mol / m^2 / s)")

xyplot(Leaf ~ (hour + 24 * doy), data = result, xlim = c(0, 24), xlab = "hour", ylab = "Leaf (kg)")
```

## Third modification: Making the `doy`, `hour`, and `time` variables consistent with a reasonable time of year

The simulation just shown takes care of making the time step
consistent with the time duration represented by the doy and hour
columns.  But the doy value of 0 corresponds to December 31, and this
is not consistent with a day having 12 hours of daylight unless we are
at an equatorial location.  We may remedy this by shifting the time
values forward by 79 days to the time of the spring equinox.  Now time
represents not just a duration in days, but more precisely, it
represents the number of (fractional) days that have elapsed since
midnight at the start of December 31 of the previous year.  Modules
that rely on this meaning for time, such as
`solar_position_michalsky`, could now be used if we wished.

```{r, realistic-time-of-year, fig.show="hold", out.width="50%"}
light_intensity <- data.frame(
    time = times/86400 + 79,
    Q = Q(times)
)

result <- run_biocro(
    initial_values,
    parameters,
    light_intensity,
    'BioCro:example_model_mass_gain',
    'BioCro:example_model_partitioning'
)


kable(result[1:4,]) # beginning
kable(result[seq(21590, 21660, 10),]) # around dawn
kable(result[43198:43202,]) # around mid day
kable(result[seq(64740, 64810, 10),]) # around dusk
kable(result[86398:86401,]) # end

xyplot(Q ~ hour, data = result, xlim = c(0, 24), xlab = "hour of the 79th day of the year\n(March 20)", ylab = "Q (mol / m^2 / s)")

xyplot(Leaf ~ hour, data = result, xlim = c(0, 24), xlab = "hour of the 79th day of the year\n(March 20)", ylab = "Leaf (kg)")
```

## References
